---
title: "FML Assignment - 5"
author: "Anushka Rampay"
date: "2024-04-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
getwd()
setwd("C:\\Users\\rampa\\Downloads")
Cereal_Data <- read.csv("Cereals.csv")
```
```{r}
head(Cereal_Data)
```

```{r}
library("cluster")
library("caret")
```
```{r}
library("dendextend")
```

```{r}
library("knitr")
library("factoextra")
```

```{r}
library("readr")
```

```{r}
Cereal_Data <- data.frame(Cereal_Data[,4:16])

#Removing Null values

Cereal_Data = na.omit(Cereal_Data)
```

```{r}
#Normalizing data

normalise_Cereals <- scale(Cereal_Data)
```

```{r}
#Use the normalized data and the Euclidean Distance method to do hierarchical clustering.

Distance <- dist(normalise_Cereals, method = "euclidean")
Hierarchical_clust <- hclust(Distance, method = "complete")
```

```{r}
# dendogram plotting

plot(Hierarchical_clust, cex = 0.7, hang = -1)
```
```{r}
#Clustering with single linkage, full linkage, and the Agnes function,average linkage and Ward.

single_Hierarchical_clust <- agnes(normalise_Cereals, method = "single")
complete_Hierarchical_clust <- agnes(normalise_Cereals, method = "complete")
average_Hierarchical_clust <- agnes(normalise_Cereals, method = "average")
ward_Hierarchical_clust <- agnes(normalise_Cereals, method = "ward")
```

```{r}
#Choosing the most efficient course of action

print(single_Hierarchical_clust$ac)

```
```{r}
print(complete_Hierarchical_clust$ac)
```
```{r}
print(average_Hierarchical_clust$ac)
```

```{r}
print(ward_Hierarchical_clust$ac)
```

With a value of 0.9046042, the ward strategy is clearly the most successful one.

***
#### Choosing the clusters
```{r}

pltree(ward_Hierarchical_clust, cex = 0.5, hang = -1, main = "Dendrogram of agnes (Using Ward)")
rect.hclust(ward_Hierarchical_clust, k = 5, border = 2:7)

```
```{r}
New_S_Group <- cutree(ward_Hierarchical_clust, k=5)
New_Data_frame_2 <- as.data.frame(cbind(normalise_Cereals,New_S_Group))
```

```{r}
#determining the stability and structure of the clusters.

fviz_cluster(list(data = New_Data_frame_2, cluster = New_S_Group))

```
```{r}
#Building Partitions

set.seed(123)
partition1<- Cereal_Data[1:50,]
partition2 <- Cereal_Data[51:74,]
```

```{r}
#Performing Hierarchical Clustering while considering k = 5.

single_sb <- agnes(scale(partition1), method = "single")
complete_sb <- agnes(scale(partition1), method = "complete")
average_sb <- agnes(scale(partition1), method = "average")
ward_sb <- agnes(scale(partition1), method = "ward")
cbind(single=single_sb$ac , complete=complete_sb$ac , average= average_sb$ac , ward= ward_sb$ac)
```

```{r}
pltree(ward_sb, cex = 0.6, hang = -1, main = "Dendogram of Agnes with Partitioned Data (Using Ward)")
rect.hclust(ward_sb, k = 5, border = 2:7)
```
```{r}
cut2 <- cutree(ward_sb, k = 5)
```

```{r}
Sbresult <- as.data.frame(cbind(partition1, cut2))
Sbresult[Sbresult$cut2==1,]
```
```{r}

centroid_one <- colMeans(Sbresult[Sbresult$cut2==1,])
Sbresult[Sbresult$cut2==2,]
```
```{r}

centroid_two <- colMeans(Sbresult[Sbresult$cut2==2,])
Sbresult[Sbresult$cut2==3,]
```
```{r}
centroid_three <- colMeans(Sbresult[Sbresult$cut2==3,])
Sbresult[Sbresult$cut2==4,]
```
```{r}

centroid_four <- colMeans(Sbresult[Sbresult$cut2==4,])
centroids <- rbind(centroid_one, centroid_two, centroid_three, centroid_four)
x3 <- as.data.frame(rbind(centroids[,-14], partition2))

```

```{r}
#figuring out the Dist.

Distance_1 <- get_dist(x3)
Matrix_1 <- as.matrix(Distance_1)
dataframe_1 <- data.frame(data=seq(1,nrow(partition2),1), Clusters = rep(0,nrow(partition2)))
for(i in 1:nrow(partition2)) 
  {dataframe_1[i,2] <- which.min(Matrix_1[i+4, 1:4])}
dataframe_1
```
```{r}
cbind(New_Data_frame_2$New_S_Group[51:74], dataframe_1$Clusters)

```

```{r}

table(New_Data_frame_2$New_S_Group[51:74] == dataframe_1$Clusters)
```

3- The elementary public schools would like to choose a set of Cereals to include in their daily cafeterias. Every day a different cereal is offered, but all Cereals should support a healthy diet. For this goal, you are requested to find a cluster of “healthy Cereals’’

```{r}
#Clustering Healthy Cereals.

Cereals_Healthy<- Cereal_Data
Cereals_Healthy_RD <- na.omit(Cereals_Healthy)
clust <- cbind(Cereals_Healthy_RD, New_S_Group)
clust[clust$New_S_Group==1,]
```

```{r}
clust[clust$New_S_Group==2,]


```
```{r}
clust[clust$New_S_Group==3,]

```
```{r}
clust[clust$New_S_Group==4,]

```
```{r}
#Mean ratings are used to select the best cluster.

mean(clust[clust$New_S_Group==1,"rating"])
```

```{r}
mean(clust[clust$New_S_Group==2,"rating"])

```

```{r}
mean(clust[clust$New_S_Group==3,"rating"])

```

```{r}
mean(clust[clust$New_S_Group==4,"rating"])

```

Based on the information above, Cluster 1 could be picked since it has the highest score.
So, Group 1 could be considered as the group for a healthy diet.

